\propertitle

### 2019-07-12 maxNt_v5.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-07-11T16:10:41+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

Answering this kind of questions always engages an element of uncertainty.
Our answers therefore involve both experimental data, such as neuronal
recordings from specific brain areas, and probabilistic assumptions --
probabilistic \enquote{models} -- about biological conditions and
mechanisms that cannot be experimentally ascertained.

Despite remarkable advances in recording technologies, the best
experimental measurements of neuronal activity can still only record a very
small sample of neurons compared to the numbers that constitute one
functionally distinguished brain region. Many probabilistic models focus on
such samples only, somehow neglecting, in their assumptions, that the
recorded neurons are a sample from a larger network. This kind of isolation
assumptions sometimes escape attention, being subtly hidden in the
mathematics. Some probabilistic models try to take also unrecorded neurons
into account, but become very complex in doing so.

In the present work we give an answer to this question: How much is our
uncertainty about the \emph{total activity} of a large neuronal network, if
we have observed the activity of a very small sample thereof? We'll first
give an answer by combining, in a straightforward way, the maximum-entropy
method and basic sampling relations of the probability calculus. Later on
we'll show that this answer can be obtained by exclusively using the
probability calculus; this point of view shows that the maximum-entropy
answer is a first approximation to the answers of \emph{four} different
questions, and also shows the assumptions lying behind it.

We apply our approach to two concrete data sets: the activity of 65 neurons
recorded from a rat's Medial Entorhinal Cortex \mynote{ref}, and the
activity of 159 neurons recorded from a macaque's Motor Cortex
\mynote{ref}\mynote{add recording length}. In the first data set, the most
interesting finding is that the most probable frequency distribution for
the total activity of the full network, if the latter is assumed to be
larger than 1\,000 neurons, has two very distinct modes, both at low
activities ($0.7\%$ and $2.6\%$ of simultaneously active neurons within
$3\;\mathrm{ms}$, see \fig***). The analogous distribution for second data
set doesn't have two modes but still presents one prominent shoulder in its
one low-activity mode. Note that these inferred features for the full
network aren't observed in the sample, nor are they inferred by the
application of maximum-entropy to the sample alone. \mynote{if we manage
  the Monte Carlo sampling or other approx: say something about the
  deviation from the most probable frequency distribution}

@@

The maximum-entropy method \citep{jaynes1957}[much clearer
in][]{jaynes1963}{sivia1996_r2006,hobsonetal1973,jaynes1985b,grandy1980}
has been used for different kinds of estimations of the neuronal activity
of various brain areas and about other phenomena of importance to the
neurosciences, for example gene and protein interaction \citep[for
example][]{martignonetal1995,bohteetal2000,shlensetal2006,schneidmanetal2006,tkaciketal2006,mackeetal2009b,tkaciketal2009,roudietal2009c,barreiroetal2010,gerwinnetal2010,mackeetal2011b,ganmoretal2011,cohenetal2011,granotatedgietal2013,mackeetal2013,tkaciketal2014b,shimazakietal2015,moraetal2015,lezonetal2006,weigtetal2009}.
This method is often used to test whether some statistics of the data, for
example second-order time correlations, is sufficient for quantifying our
uncertainty about the data. It can be considered as an approximation of
probabilistic models based on various assumptions of inferential
sufficiency \citep{jaynes1986d_r1996,portamana2017} -- for example of
empirical moments or of long-run frequencies.

@@

@@

The context we consider is as follows. During an experimental session we
have recorded the spiking activities of $n$ neurons for a certain amount of
time. These neurons are our \enquote{sample} or \enquote{subnetwork}. Their
spikes are binned into $T$ time bins and binarized to $\set{0,1}$ values in
each bin. Call $\ya_{t}$ the number of neurons that fire during time bin
$t$, divided by the total number of neurons $n$; this is the
\emph{normalized total activity} of the sample, or just \enquote{activity}
for short. It is also the network-averaged activity of the neurons.
Obviously $\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if $\ya_{t}=0$, no
neuron spikes during bin $t$; if $\ya_{t}=1$, all spike at some point
during bin $t$. For brevity, let's say \enquote{at $t$} for \enquote{during
  time bin $t$}. From the activities $\set{\ya_{t}}$ we can count how many
times the activity levels $\ya=0$, $\ya=1/n$, and so on appeared during the
recording, obtaining the measured relative-frequency distribution
$(\yff_{\ya}) \defs \yf$. We can also consider the sample activity at time
bins \emph{outside} of the recorded range. Such activity is unknown to us,
of course.

\mynote{maybe move this paragraph to the intro?}Present-day technologies
enable the recording of neuronal activity from small brain areas
\mynote{be more specific}. For many animal species, the neurons that are
recorded within the area are not -- and at present cannot be --
specifically chosen from among the rest, owing to several limiting factors;
for example, limitations in the precision with which electrodes are
inserted, or with which neurons are targeted by viruses. In fact, the set
of recorded neurons may even change during very long recordings or across
experimental sessions. We assume that there's a area, comprising a
network of $N$ neurons, from which other sets of neurons had or will have
the same probability of being recorded, even in other experimental
sessions, as the set of $n$ neurons that was actually recorded in this
session. This is our \enquote{full network}\mynote{how about calling it
  \enquote{the pool}?}. The normalized total activity of these $N$ neurons
at $t$ is $\yA_{t}$. The relative frequencies of the various activity
levels during the recording were $(\yFF_{\yA}) \defs \yF$. We don't know the
values $\yA_{t}$ at each $t$, or the frequency distribution $\yF$. We only
know for certain that $\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that
$N\yA_{t} \ge n\ya_{t}$ for obvious reasons. For the time being we assume
that we know $N$; in \sect*** we discuss the consequences of our lack of
precise knowledge about this number.

Our questions concern general features of the total activity $\yA$ of the
full network during and after the recording, and across sessions under the
same study conditions. For example: what was its frequency distribution
during the recording? How much does this frequency distribution change
across sessions? How much total activity should we expect at any time bin
during a recording? We cannot answer these questions with certainty; we can
only give distributions of probability over their possible answers. The
approach presented here gives such probability distributions.

\mynote{Move this to intro?}We want to stress the usefulness of making
quantified guesses about the full-network activity. First of all, this seems to be
the primary idea behind recording a sample. Second, it allows us to make
comparisons across experimental sessions; such comparisons would me
difficult or meaningless made at the level of recorded samples, which
generally comprise non-overlapping sets of neurons differing also in size.

@@

Regarding~\ref{item:maxent_step}, we assume that you're familiar with the
maximum-entropy method. In short, it amounts to a pair of prescriptions:
choose the distribution having maximum Shannon entropy among those
satisfying specific convex constraints, such as fixed averages; and judge
these averages to be equal to measured averages. We add two remarks about
this method that are seldom made in the literature. First, the distribution
$\px(\yA)$ given by this method is the zeroth-order approximation \citep[in
the sense of Laplace's
method:][\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000} of
four different distributions for the full network:
\begin{itemize}
\item the most probable \emph{frequency} distribution for the total
  activity across the \emph{recorded} bins, %$\yFF(\yA)$,
\item the most probable \emph{frequency} distribution for the total
  activity in a very long run of \emph{new} time bins, %$\ynuu(\yA)$,
\item the \emph{probability} distribution for the value of the total
  activity at any time bin among those \emph{recorded},
  %$\yp(\yA_{t} \| \yH)$ with $1\le t \le T$,
\item the \emph{probability} distribution for the value of the total
  activity at a \emph{new} time bin.
  % $\yp(\yA_{t} \| \yH)$ with $t<1$ or $t>T$
\end{itemize}
The maximum-entropy distribution is thus an approximation of our
uncertainty about four completely different quantities. Note that the four
distributions above numerically differ in higher-order approximations. We
discuss their differences in \sect***. Second, the maximum-entropy method
based on the Shannon entropy implicitly makes some assumptions about the
probabilities for the long-run frequency distributions
\citep{jaynes1986d_r1996,portamana2009,portamana2017}. We discuss these
assumptions in \sect***.

@@

The number of moments used with this method depends on the questions and
hypotheses that a researcher is exploring; for example, hypotheses of
sufficient statistics, such as the sufficiency of pairwise correlations to
quantify our uncertainty about network activity. In the present work we
only want to introduce the general method without entering into biological
questions of this kind.

The particular case in which $n$ moments are constrained is especially
important: it corresponds to fully constraining the marginal frequency
distribution for the activity of the sample neurons. In this case, our
uncertainty about the full-network activity is based on all available
measured frequency data. Note that application of the maximum-entropy
method \emph{at the sample level} is trivial and meaningless in this case
-- it just gives back the measured frequency distribution. But application
of the method \emph{at the level of the full network} is not trivial.

@@

Consider the question: What is the relative-frequency distribution for the
different activity levels for the full-network, across the recorded bins?
Call this distribution $(\yFF_{\yA}) \defs \yF$: activity level $\yA$
appeared in $T\times\yFF(\yA)$ out of $T$ time bins. We don't know $\yF$,
and our uncertainty about its possible values is a probability distribution
$\pf(\yF\|\yD\yI)$ conditional on data $\yD$ and prior knowledge $\yI$,
which we approximate with a continuous density
\begin{equation}
  \label{eq:density_F}
  \pf(\yF \| \yD \yI)\;\di\yF.
\end{equation}
In appendix*** we give a summary of how this density is derived. Note that
this is a probability distribution over frequency distributions, so we must
be careful with terminology to avoid confusion.

Now consider the related question: What is the activity $\yA_{t}$ of the
full network at some specific time bin $t$? If we don't know the exact
sequence of activities during the recording then we don't know $\yA_{t}$;
if we know the frequency distribution $\yF$ of the activity levels, then
our probability for $\yA_{t}$ is $\yFF_{\yA_{t}}$, out of combinatorial
reasons. But if we are also uncertain about $\yF$, with
probability~\eqref{eq:density_F}, then our probability for $\yA_{t}$ by
marginalization is
\begin{equation}
  \label{eq:probability_AT}
  \pf(\yA_{t} \| \yD \yI) =
  \int \yFF_{\yA_{t}}\;   \pf(\yF \| \yD \yI)\;\di\yF;
\end{equation}
our probability distribution for $\yA_{t}$ is thus simply the mean
$\int \yF\, \pf(\yF \| \yD \yI)\,\di\yF$.

@@

\textbullet\ The density
\begin{equation}
\exp[-\yL\;\ysh(\yF;\yR)]\label{eq:entropic_prior}
\end{equation}
expresses our pre-data
uncertainty about the possible frequency distributions. Here
$\ysh(\yF;\yR)\defd \sum_{\yA}\yFF_{\yA}\ln(\yFF_{\yA}/\yRR_{\yA})$ is the
relative entropy or discrimination information
\citep{kullback1987,jaynes1963,hobson1969,hobsonetal1973}, $\yR$ is a
reference distribution, and $\yL$ a positive parameter. This density
expresses a belief proportional to the number of ways in which the
frequency distribution $\yF$ can be realized if we assume that the
activities appear with frequencies $\yR$ in the long run, as can be seen
using Stirling's approximation. For $\yR$ we choose the uniform
distribution, but any biologically sensible distribution, such as one
decreasing exponentially with the activity, leads to the same final
results. The parameter $\yL$ roughly represents the number of data points
necessary to change our initial belief, and we set it equal to $50$. This
density is similar to an \enquote{entropic prior}
\citep{rodriguez1991,skilling1998,catichaetal2004,portamana2017}[see
esp.][\sect~3]{neumann2007}.

@@

\textbullet\ The normal density
$\exp\Bigl(-\frac{\abs{\ycm\yF - \yc}^{2}}{2\ycs^{2}}\Bigr)$ is the
likelihood of the frequency $\yF$ in view of our data. The matrix $\ycm$
expresses the linear constraints on $\yF$, and the tuple $\yc$ the measured
values of such constraints. For example, if we consider the full frequency
distribution $\yf$ of the sample, then $\ycm$ effects the marginalization
to a sample of size $n$ through a hypergeometric distribution (sampling
without replacement):
\begin{equation}
  \label{eq:constraint_hypergeometric}
\ycmm_{\ya,\yA} \defd
  \binom{n}{n\ya}\binom{\yNv-n}{\yNv \yA-n\ya}\binom{\yNv}{\yNv \yA}^{-1}
\end{equation}
and $\yc \defd \yf$ are the measured frequencies. If we only consider some
factorial moments, instead, then
\begin{equation}
  \label{eq:constraint_hypergeometric}
  \ycmm_{m,\yA} \defd
  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}
\end{equation}
as for \eqn~\eqref{eq:equal_m_factorialmoments}, and $\yc$ is the tuple of
measured factorial moments. The standard deviations $\ycs$ reflect our
uncertainty about the measurement results, which can come for example from
the spike-sorting procedure. Their values are discussed in appendix***.

@@

The post-data belief density~\eqref{eq:density_proportional} has a
maximum-entropy distribution as a mode approximation when $\ycs$ is small,
$\yL$ is large, and $\ycs L$ is small. Its mode in fact will be the
constrained $\yF$ that maximizes the relative entropy $\ysh(\yF;\yR)$. It
should be noted that a pre-data belief different from the entropic
one~\eqref{eq:entropic_prior} will lead to a different mode approximation.
For example, a Dirichlet density would lead to a maximum-entropy
distribution with \emph{Burg's} entropy \citep{burg1975} instead of
Shannon's \citep{jaynes1986d_r1996,portamana2009}.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-07-12T11:25:44+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

Answering this kind of questions always engages an element of uncertainty.
Our answers therefore involve experimental data, such as neuronal
recordings from specific brain areas, and probabilities or \dobs, based on
prior knowledge, about biological conditions and mechanisms that cannot be
experimentally ascertained. Such \dobs\ are often formulated as simplified
\enquote{models} to be mathematically more tractable.

Despite remarkable advances in recording technologies, the best
experimental measurements of neuronal activity can still only record a very
small sample of neurons compared to the numbers that constitute a
functionally distinguished brain region. Many probabilistic models focus on
such samples only, somehow neglecting, in their assumptions, that the
recorded neurons are a sample from a larger network. This kind of isolation
assumptions sometimes escape attention, being subtly hidden in the
mathematics. Some probabilistic models try to take also unrecorded neurons
into account, but become very complex in doing so.

In the present work we give an answer to this question: How much can the
total activity of a large neuronal network be, if we have observed the
activity of a very small sample thereof? We'll quantify our \dobs\ about
the possible answers by combining, in a straightforward way, the
maximum-entropy method and basic sampling relations of the probability
calculus. Later on we'll show that our \dobs\ can be quantified by
exclusively using the probability calculus. This derivation will provide a
more accurate quantification, revealing that the maximum-entropy answer is
only a first approximation.

We apply our approach to two concrete data sets: the activity of 65 neurons
recorded from a rat's Medial Entorhinal Cortex \mynote{ref}, and the
activity of 159 neurons recorded from a macaque's Motor Cortex
\mynote{ref}\mynote{add recording length}. In the first data set, the most
interesting finding is that the most probable frequency distribution for
the total activity of the full network has two very distinct modes, both at
low activities, %($0.7\%$ and $2.6\%$ of simultaneously active neurons
% within $3\;\mathrm{ms}$,
see \fig~\ref{fig:plots_6mom}. The analogous frequency distribution for the
second data set doesn't have two modes but still presents one prominent
shoulder in its one low-activity mode. Note that these guessed features of
the full network aren't observed in the sample, nor can they be inferred by
the application of maximum-entropy \emph{to the sample alone}.
\mynote{Monte Carlo sampling: say something about the deviation from the
  most probable frequency distribution}

@@

The maximum-entropy or minimum-relative-entropy method
\citep{jaynes1957}[much clearer
in][]{jaynes1963}{sivia1996_r2006,hobsonetal1973,jaynes1985b,grandy1980}
has been used for different kinds of estimations of the neuronal activity
of various brain areas and about other phenomena of importance to the
neurosciences, for example gene and protein interaction \citep[for
example][]{martignonetal1995,bohteetal2000,shlensetal2006,schneidmanetal2006,tkaciketal2006,mackeetal2009b,tkaciketal2009,roudietal2009c,barreiroetal2010,gerwinnetal2010,mackeetal2011b,ganmoretal2011,cohenetal2011,granotatedgietal2013,mackeetal2013,tkaciketal2014b,shimazakietal2015,moraetal2015,lezonetal2006,weigtetal2009}.
This method is often used to test whether some statistics of the data, for
example second-order time correlations, is sufficient for quantifying our
\dob\ about some quantities of the system. It can be considered as an
approximation of probabilistic models based on various assumptions of
inferential sufficiency \citep{jaynes1986d_r1996,portamana2017}.

@@

Our notation and terminology follow \textsc{iso} standards
\citep{iso1993,iso2006,iso2006b} and Jaynes \citey{jaynes1994_r2003} for
\dobs. We often simply say \enquote{belief} for \enquote{\dob}.

@@

The context we consider is as follows. During an experimental session we
have recorded the spiking activities of $n$ neurons for a certain amount of
time. These neurons are our \enquote{sample} or \enquote{subnetwork}. Their
spikes are binned into $T$ time bins and binarized to $\set{0,1}$ values in
each bin. Call $\ya_{t}$ the number of neurons that fire during time bin
$t$, divided by the total number of neurons $n$; this is the
\emph{normalized total activity} of the sample, or just \enquote{activity}
for short. It is also the network-averaged activity of the neurons.
Obviously $\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if $\ya_{t}=0$, no
neuron spikes during bin $t$; if $\ya_{t}=1$, all spike at some point
during bin $t$. For brevity, let's say \enquote{at $t$} for \enquote{during
  time bin $t$}. From the activities $\set{\ya_{t}}$ we can count how often
the activity levels $\ya=0$, $\ya=1/n$, and so on appeared during the
recording, obtaining the distribution of measured relative frequencies
$(\yff_{\ya}) \defs \yf$. We can also consider the sample activity at time
bins \emph{outside} of the recorded range. Such activity is unknown to us,
of course.

\mynote{maybe move this paragraph to the intro?}Present-day technologies
enable the recording of neuronal activity from small brain areas
\mynote{be more specific}. For many animal species, the neurons that are
recorded within the area are not -- and at present cannot be --
specifically chosen from among the rest, owing to several limiting factors;
for example, limitations in the precision with which electrodes are
inserted, or with which neurons are targeted by viruses. In fact, the set
of recorded neurons may even change during very long recordings or across
experimental sessions. We assume that there's a area, comprising a
network of $N$ neurons, from which other sets of neurons had or will have
the same probability of being recorded, even in other experimental
sessions, as the set of $n$ neurons that was actually recorded in this
session. This is our \enquote{full network}.\mynote{how about calling it
  \enquote{the pool}?} The normalized total activity of these $N$ neurons
at $t$ is $\yA_{t}$. The relative frequencies of the various activity
levels during the recording were $(\yFF_{\yA}) \defs \yF$. We don't know the
values $\yA_{t}$ at each $t$, or the frequency distribution $\yF$. We only
know for certain that $\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that
$N\yA_{t} \ge n\ya_{t}$ for obvious reasons. For the time being we assume
that we know $N$; in \sect*** we discuss the consequences of our lack of
precise knowledge about this number.

Our questions concern general features of the total activity $\yA$ of the
full network during and after the recording, and across sessions under the
same study conditions. For example: what was its frequency distribution
during the recording? How much does this frequency distribution change
across sessions? How much total activity should we expect at any time bin
during a recording? We cannot answer these questions with certainty; we can
only give distributions of probability or \dobs\ over their possible answers. The
approach presented here gives such probability distributions.

\mynote{Move this to intro?}We want to stress the usefulness of making
quantified guesses about the full-network activity. First of all, this
seems to be the primary idea behind recording a sample. Second, it allows
us to make comparisons across experimental sessions; such comparisons would
me difficult or meaningless if made with the recorded samples, which
generally comprise non-overlapping sets of neurons and differ in size.

@@

Regarding~\ref{item:maxent_step}, we assume that you're familiar with the
maximum-entropy method. We actually use the \emph{minimum-relative-entropy}
method \citep{hobsonetal1973}, but call it \enquote{maximum-entropy} for
brevity. It amounts to a pair of prescriptions: choose the distribution,
among those satisfying specific convex constraints, such as fixed
expectations, that has minimum relative entropy with respect to a reference
distribution, often taken to be the uniform one; and judge those expectations
to be equal to measured averages. We add two remarks about this method that
are seldom made in the literature. First, the distribution $\px(\yA)$ given
by this method is the zeroth-order approximation \citep[in the sense of
Laplace's
method:][\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000} of
four different distributions for the full network:
\begin{itemize}
\item the most probable \emph{frequency} distribution for the total
  activity across the \emph{recorded} bins, %$\yFF(\yA)$,
\item the \emph{belief} distribution for the value of the total
  activity at any time bin among those \emph{recorded},
  %$\yp(\yA_{t} \| \yH)$ with $1\le t \le T$,
\item the most probable \emph{frequency} distribution for the total
  activity in a very long run of \emph{new} time bins, %$\ynuu(\yA)$,
\item the \emph{belief} distribution for the value of the total
  activity at a \emph{new} time bin.
  % $\yp(\yA_{t} \| \yH)$ with $t<1$ or $t>T$
\end{itemize}
The maximum-entropy distribution is thus an approximation of our belief
distribution about four completely different quantities. Note that the four
distributions above numerically differ in higher-order approximations. We
discuss their differences in \sect***. Second, the maximum-entropy method
based on the Shannon entropy implicitly makes some assumptions about the
probabilities for the long-run frequency distributions
\citep{jaynes1986d_r1996,portamana2009,portamana2017}. We discuss these
assumptions in \sect***.

@@

The number of moments used with this method depends on the questions and
hypotheses that a researcher is exploring; for example, hypotheses of
sufficient statistics, such as the sufficiency of pairwise correlations to
quantify our \dob\ about network activity. In the present work we
only want to introduce the general method without entering into biological
questions of this kind.

The particular case in which $n$ moments are constrained is especially
important: it corresponds to fully constraining the marginal frequency
distribution for the activity of the sample neurons. In this case, our
belief about the full-network activity is based on all available
measured frequency data. Note that application of the maximum-entropy
method \emph{at the sample level} is trivial and meaningless in this case
-- it just gives back the measured frequency distribution. But application
of the method \emph{at the level of the full network} is not trivial.

@@

Consider the question: What is the relative-frequency distribution for the
different activity levels for the full-network, across the recorded bins?
Call this distribution $(\yFF_{\yA}) \defs \yF$: activity level $\yA$
appeared in $T\times\yFF(\yA)$ out of $T$ time bins. We don't know $\yF$,
and our beliefs about its possible values have a distribution
$\pf(\yF\|\yD\yI)$ conditional on data $\yD$ and prior knowledge $\yI$,
which we approximate with a continuous distribution
\begin{equation}
  \label{eq:density_F}
  \pf(\yF \| \yD \yI)\;\di\yF.
\end{equation}
In appendix*** we give a summary of how this distribution is derived. Note that
this is a probability distribution over frequency distributions, so we must
be careful with terminology to avoid confusion.

Now consider the related question: What is the activity $\yA_{t}$ of the
full network at some specific time bin $t$? If we don't know the exact
sequence of activities during the recording then we don't know $\yA_{t}$;
if we know the frequency distribution $\yF$ of the activity levels, then
our \dob\ about $\yA_{t}$ is $\yFF_{\yA_{t}}$, out of combinatorial
reasons. But if we are also uncertain about $\yF$, with
belief distribution~\eqref{eq:density_F}, then our \dob\ about $\yA_{t}$ by
marginalization is
\begin{equation}
  \label{eq:probability_AT}
  \pf(\yA_{t} \| \yD \yI) =
  \int \yFF_{\yA_{t}}\;   \pf(\yF \| \yD \yI)\;\di\yF;
\end{equation}
our belief distribution for $\yA_{t}$ is thus simply the mean
$\int \yF\, \pf(\yF \| \yD \yI)\,\di\yF$.

@@

\textbullet\ The density
\begin{equation}
\exp[-\yL\;\ysh(\yF;\yR)]\label{eq:entropic_prior}
\end{equation}
expresses our pre-data
beliefs about the possible frequency distributions. Here
$\ysh(\yF;\yR)\defd \sum_{\yA}\yFF_{\yA}\ln(\yFF_{\yA}/\yRR_{\yA})$ is the
relative entropy or discrimination information
\citep{kullback1987,jaynes1963,hobson1969,hobsonetal1973}, $\yR$ is a
reference distribution, and $\yL$ a positive parameter. This density
expresses a belief proportional to the number of ways in which the
frequency distribution $\yF$ can be realized if we assume that the
activities appear with frequencies $\yR$ in the long run, as can be seen
using Stirling's approximation. For $\yR$ we choose the uniform
distribution, but any biologically sensible distribution, such as one
decreasing exponentially with the activity, leads to the same final
results. The parameter $\yL$ roughly represents the number of data points
necessary to change our initial belief, and we set it equal to $50$. This
density is similar to an \enquote{entropic prior}
\citep{rodriguez1991,skilling1998,catichaetal2004,portamana2017}[see
esp.][\sect~3]{neumann2007}.

@@

\textbullet\ The normal density
$\exp\Bigl(-\frac{\abs{\ycm\yF - \yc}^{2}}{2\ycs^{2}}\Bigr)$ is the
likelihood of the frequency $\yF$ in view of our data. The matrix $\ycm$
expresses the linear constraints on $\yF$, and the tuple $\yc$ the measured
values of such constraints. For example, if we consider the full frequency
distribution $\yf$ of the sample, then $\ycm$ effects the marginalization
to a sample of size $n$ through a hypergeometric distribution (sampling
without replacement):
\begin{equation}
  \label{eq:constraint_hypergeometric}
\ycmm_{\ya,\yA} \defd
  \binom{n}{n\ya}\binom{\yNv-n}{\yNv \yA-n\ya}\binom{\yNv}{\yNv \yA}^{-1}
\end{equation}
and $\yc \defd \yf$ are the measured frequencies. If we only consider some
factorial moments, instead, then
\begin{equation}
  \label{eq:constraint_hypergeometric}
  \ycmm_{m,\yA} \defd
  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}
\end{equation}
as for \eqn~\eqref{eq:equal_m_factorialmoments}, and $\yc$ is the tuple of
measured factorial moments. The standard deviations $\ycs$ reflect our
beliefs about the measurement results, which can come for example from
the spike-sorting procedure. Their values are discussed in appendix***.

@@

The post-data belief density~\eqref{eq:density_proportional} has a
maximum-entropy distribution as a mode approximation when $\ycs$ is small,
$\yL$ is large, and $\ycs L$ is small: the mode will be the constrained
$\yF$ that minimizes the relative entropy $\ysh(\yF;\yR)$. It should be
noted that a pre-data belief different from the entropic
one~\eqref{eq:entropic_prior} will lead to a different mode approximation.
For example, a Dirichlet density would lead to a maximum-entropy
distribution with \emph{Burg's} entropy \citep{burg1975} instead of
Shannon's \citep{jaynes1986d_r1996,portamana2009}.
